Optimized average-based embeddings for word sense disambiguation problem

### 2019-02-11 paper/main.tex
Use centering on figures
Improve baseline model description

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[outdir=./]{epstopdf}
\usepackage{bm}

@@

\begin{figure}
    \label{fig:keywords}
    \caption{Ambiguous words with their meanings (keywords) from the dataset}
    \includegraphics[scale=0.35]{res/keywords.png}
\end{figure}

@@

\subsection{Limitations}
We are aware that our method has some limitations.
First of all, it may be impossible to achieve the optimal solution, as we can only optimize keyword embeddings, leaving context embeddings fixed in the multidimensional space.
Therefore, it is possible that contexts for specific keyword overlap on contexts for other keyword.

@@

We compare our results with two metrics: top-1 accuracy, where we are only interested, if the correct keyword is the closest one to a specific context describing it, and top-3 accuracy, where we check if we are in the closest three keywords.

\subsection{Baseline}
First of experiments is to check how out-of-the-box model works without extra optimizations. We perform tests with different context lengths, as a score we choose top 1 and top 3 accuracy. Results are shown in the Fig. \ref{fig:baseline}.

\begin{figure}
    \label{fig:baseline}
    \caption{Top accuracy of base model with different context length}
    \includegraphics[scale=0.5]{res/baseline_context_top_acc.png}
\end{figure}

We see that the base model exceeds 90 percentages of top 3 accuracy for very short context length. It may be caused by low number of meanings of each ambiguous word.
It can also be observed that top 1 score is the best for context length of 3 and therefore we decided to use it as a baseline for next two experiments.

\subsection{Context length optimization}
Once we have base model results we check how simple alpha optimization improves accuracy based on context length. To do this, we choose arbitrarily a value for alpha parameter and run tests after one epoch of training the model. Since we average results from 30 runs we mark the measurement uncertainty on the graph. Results are shown in the Fig. \ref{fig:exp1}.
\begin{figure}
    \label{fig:exp1}
    \caption{Comparison of the base model and a little bit optimized model based on different context length}
    \includegraphics[scale=0.6]{res/exp1_context_top_acc.png}
\end{figure}
Independent of some noise caused by random order of training data, the optimized model is much better than the base model (12 - 36\% of improvement). Interestingly, the improvement is most visible for long context lengths, which achieved relatively poor results in the baseline experiment. It may be caused by huge value of alpha parameter. It can improve bad baseline prediction but when baseline prediction is quite good, the improvement is worse because update step is too long.

@@

\begin{figure}
    \label{fig:exp2}
    \caption{Based on fixed context length, comparison of the base model and a model optimized with various alpha parameters. }
    \includegraphics[scale=0.6]{res/exp2_alpha_top_acc.png}
\end{figure}

@@

\begin{figure}
    \label{fig:exp3}
    \caption{Based on fixed context length and alpha parameter value, comparison of the alpha-optimized model and a alpha-optimized model with various beta parameters. }
    \includegraphics[scale=0.5]{res/exp3_beta_top_acc.png}
\end{figure}
Beta parameter is also a scalar value and it is responsible for increasing distance between keyword and incorrect context (Equation \ref{eq:beta_optimization}) . In the graph of results we see that for low value of beta there is an improvement in top 1 accuracy, for top 3 accuracy the score goes under baseline value. It is interesting as we improve top 1 score and worsen top 3 score. Probably beta parameter should not be optimized during single epoch training and should be very slow as increasing distances between two incorrect embeddings may cause that we also increase distance between correct pair of embeddings.

\subsection{Longer training}
Previous optimization are applied for only one epoch of training. The goal of this experiment is to check how number of epochs affects final accuracy. We run up to 5 epochs of training using parameters obtained in previous experiments. Results are shown in the Fig. \ref{fig:exp4_2}.
\begin{figure}
    \label{fig:exp4_1}
    \caption{5 epoch training of optimized model}
    \includegraphics[scale=0.6]{res/exp4_epoch_top_acc.png}
\end{figure}
We see that there is almost no improvement after first epoch of training. Probably the problem is that alpha and beta parameters are too big. We also check what happens when we choose lower values of alpha/beta parameters. Results are shown in the Fig. \ref{fig:exp4_2}.
\begin{figure}
    \label{fig:exp4_2}
    \caption{Comparison of multiple optimized models across 5 epochs of training. }
    \includegraphics[scale=0.6]{res/exp4_epoch_multiconfiguration_top_acc.png}
\end{figure}
Longer training shows that optimal parameters obtained from previous experiments are not good for longer training and lower values of alpha and beta parameters are better. Probably we should use additional decays parameters to decrease values of alpha and beta parameters.

>>>>

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[outdir=./]{epstopdf}
\usepackage{bm}
\usepackage{float}

@@

\begin{figure}
    \centering
    \label{fig:keywords}
    \caption{Ambiguous words with their meanings (keywords) from the dataset}
    \includegraphics[scale=0.35]{res/keywords.png}
\end{figure}

@@

\subsection{Limitations}
\label{sec:limitations}
We are aware that our method has some limitations.
First of all, it may be impossible to achieve the optimal solution, as we can only optimize keyword embeddings, leaving context embeddings fixed in the multidimensional space.
Therefore, it is possible that contexts for specific keyword overlap on contexts for other keyword.

@@

We compare results on the test set with two metrics: top-1 accuracy, where we are only interested, if the correct keyword is the closest one to a specific context describing it, and top-3 accuracy, where we check if we are in the closest three keywords.

\subsection{Baseline}
For the baseline model, we check how out-of-the-box keyword (\ref{sec:kw_embed}) and context embeddings (\ref{sec:ctx_embed}) work without any extra optimizations.
We evaluate the performance of the model with different context lengths, to see how it affects top-1 and top-3 accuracies.
Results are shown in the Fig. \ref{fig:baseline}.

\begin{figure}[H]
    \centering
    \label{fig:baseline}
    \caption{Top accuracy of base model with different context length}
    \includegraphics[scale=0.5]{res/baseline_context_top_acc.png}
\end{figure}

We can see, that the model does relatively well, even unoptimized.
Top-3 accuracy is about 85-90\%, which is probably caused by a low number of meanings for each ambiguous word.
Top-1 accuracy for shorter context lengths can go as high as 65\% but decreases with longer contexts.
As suspected (\ref{sec:limitations}), this is most likely due to the fact, that the average of many word embeddings may make some contexts similar to each other, therefore making it harder to distinguish between some meanings.

The best result w.r.t. top-1 accuracy was achieved with \(l=3\), which is why we choose this context length as a starting point for next experiments.

\subsection{Context length optimization}
Once we have base model results we check how simple alpha optimization improves accuracy based on context length. To do this, we choose arbitrarily a value for alpha parameter and run tests after one epoch of training the model. Since we average results from 30 runs we mark the measurement uncertainty on the graph. Results are shown in the Fig. \ref{fig:exp1}.
\begin{figure}
    \centering
    \label{fig:exp1}
    \caption{Comparison of the base model and a little bit optimized model based on different context length}
    \includegraphics[scale=0.6]{res/exp1_context_top_acc.png}
\end{figure}
Independent of some noise caused by random order of training data, the optimized model is much better than the base model (12 - 36\% of improvement). Interestingly, the improvement is most visible for long context lengths, which achieved relatively poor results in the baseline experiment. It may be caused by huge value of alpha parameter. It can improve bad baseline prediction but when baseline prediction is quite good, the improvement is worse because update step is too long.

@@

\begin{figure}
    \centering
    \label{fig:exp2}
    \caption{Based on fixed context length, comparison of the base model and a model optimized with various alpha parameters. }
    \includegraphics[scale=0.6]{res/exp2_alpha_top_acc.png}
\end{figure}

@@

\begin{figure}
    \centering
    \label{fig:exp3}
    \caption{Based on fixed context length and alpha parameter value, comparison of the alpha-optimized model and a alpha-optimized model with various beta parameters. }
    \includegraphics[scale=0.6]{res/exp3_beta_top_acc.png}
\end{figure}
Beta parameter is also a scalar value and it is responsible for increasing distance between keyword and incorrect context (Equation \ref{eq:beta_optimization}) . In the graph of results we see that for low value of beta there is an improvement in top 1 accuracy, for top 3 accuracy the score goes under baseline value. It is interesting as we improve top 1 score and worsen top 3 score. Probably beta parameter should not be optimized during single epoch training and should be very slow as increasing distances between two incorrect embeddings may cause that we also increase distance between correct pair of embeddings.

\subsection{Longer training}
Previous optimization are applied for only one epoch of training. The goal of this experiment is to check how number of epochs affects final accuracy. We run up to 5 epochs of training using parameters obtained in previous experiments. Results are shown in the Fig. \ref{fig:exp4_2}.
\begin{figure}
    \centering
    \label{fig:exp4_1}
    \caption{5 epoch training of optimized model}
    \includegraphics[scale=0.6]{res/exp4_epoch_top_acc.png}
\end{figure}
We see that there is almost no improvement after first epoch of training. Probably the problem is that alpha and beta parameters are too big. We also check what happens when we choose lower values of alpha/beta parameters. Results are shown in the Fig. \ref{fig:exp4_2}.
\begin{figure}
    \centering
    \label{fig:exp4_2}
    \caption{Comparison of multiple optimized models across 5 epochs of training. }
    \includegraphics[scale=0.6]{res/exp4_epoch_multiconfiguration_top_acc.png}
\end{figure}
Longer training shows that optimal parameters obtained from previous experiments are not good for longer training and lower values of alpha and beta parameters are better. Probably we should use additional decays parameters to decrease values of alpha and beta parameters.
